---
title: "beepboop2"
author: 'Margaret Hammond'
output: 
  html_document:
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(purrr)
library(minpack.lm)
library(broom)
library(patchwork)
library(kableExtra)
library(janitor)
library(modelr)
```

## 1. Overview
In this task, I will be using non linear least squares to estimate parameters of a length to weight model for lizard populations in New Mexico. 
PUT OVERVIEW HERE

## Explore the Data

### 2.1 Read in the data
```{r}
lizards <- read_csv(here("data", "lizard.csv")) %>% 
  mutate(log_length = log(SV_length), # use log() to linearize the model
         log_weight = log(weight))
```
## Set up the model

Must set up the model to be able to make guesses for coefficients a and b. Take the data and create a linear model in order for OLS regression to occur.


$$
W = a(SVL)^b
$$
In this model, W is weight, and SVL is snout to vent length.


```{r}
lizards_model <- function(a, SVL, b) {
  W = a*SVL^b
  return(W)
}
```



```{r}
parameter_guess <- lm(log_weight ~ log_length, data = lizards) # Created for OLS

lizards_coef <- coef(parameter_guess) 

lizards_coef
```


## Functions



```{r}

lizards_nls <- nls(weight ~ lizards_model(a, SV_length, b),
                   data = lizards,
                   start = list(
                     a = 2.718^(lizards_coef[1]/lizards_coef[2]),
                     b = lizards_coef[2]),
                   trace = TRUE)
```

### Table of Model
This table shows the estimated parameters for the nls model. 

```{r}
tidy_nls <- tidy(lizards_nls)

tidy_nls$p.value <- ifelse(tidy_nls$p.value < .001, paste('< .001'))

tidy_nls %>% 
  select(-statistic) %>% 
  kable(col.names = c('Parameter', 
                      'Estimated', 
                      'Standard Error', 
                      'P Values'), 
        caption = 'Table 1. Parameter estimates of coefficient a and b for the nls model.',
        digits = 4) %>% 
  kable_classic(bootstrap_options = "striped", full_width = FALSE)
```


### NLS Predictions

I will use the nls model to generate predictions from the data. **present fitted model on a plot with female and male lizards separated by color. INCLUDE THE MODEL IN KABLE OUTPUT OF THE HTML**

```{r}
general_mdl_pred <- predict(lizards_nls)


lizards_complete <- cbind(lizards, general_mdl_pred)

ggplot(data = lizards_complete,
       aes(x = SV_length, 
           y = general_mdl_pred)) +
  geom_point(aes(x = SV_length, y = weight, color = sex)) +
  scale_color_manual(values = c('pink', 'blue2')) +
  geom_line(color = 'indianred2', size = 1) +
  labs(x = 'Snout to Vent Length (mm)',
       y = 'Weight (g)',
       title = "Lizard NLS Model Predictions",
       caption = 'Figure 1. NLS vs measured weight and
       snout to vent length of lizards in
       Jornada Basin LTER site, 1989-2006') +
  theme_minimal()

```


### Fit NLS to Western Whiptail lizard
Filter out the dataset for male Western Whiptail lizard (Cnemidophorus tigrisatus). Fit an nls model to this subset. Compare the output from the species specific nls model to the general nls model for all species by graphing the model fits on the Western Whiptail male data. Report the RMSE for both models in the figure caption. Also in the figure caption briefly discuss which model should be used and why.


```{r}

# filter out just male Western Whiptail

whip_sub <- lizards %>% 
  filter(spp == 'CNTI', 
         sex == 'M') %>% 
  mutate(log_weight = log(weight)) %>% 
  mutate(log_length = log(SV_length))


whip_pred <- lm(log_weight ~ log_length, data = whip_sub)

whip_coef <- coefficients(whip_pred)

whip_nls <- nls(weight ~ lizards_model(a, SV_length, b),
                   data = whip_sub,
                   start = list(
                     a = 2.718^(whip_coef[1]/whip_coef[2]),
                     b = whip_coef[2]),
                   trace = TRUE)

tidy_whip <- tidy(whip_nls)

tidy_whip

```
**PROLLY INLINE CODE HERE**

### Compare the Species NLS to General NLS

```{r}
lizards_general_whip <- lizards_complete %>% 
  filter(spp == 'CNTI',
         sex == 'M')

whip_mdl_pred <- predict(whip_nls)

complete_pred <- cbind(lizards_general_whip, whip_mdl_pred)

rmse_whip <- rmse(whip_nls, complete_pred)

rmse_gen <- rmse(lizards_nls, complete_pred)

rmse_complete <- cbind(rmse_whip, rmse_gen)
rmse_complete # rmse_whip has a lower value than rmse_gen, RMSE of whip specific set is better

```

```{r}
# Create plot showing the differences between the general NLS and species specific nls for Western Whip
caption = 'Figure 2. NLS species specific model for Western Whiptail (in red) vs NLS of all lizard species (in navy) in Jornada Basin LTER site, 1989 - 2006. Black points represent actual Western Whiptail measurements of weight (g) to snout to vent length (mm). RMSE calculations done for both models, species specific model contained a lower RMSE score of 3.349 compared to general NLS model RMSE of 3.562. The species specific RMSE should be used when doing RMSE calculations for Western Whiptail, it is better fit for the data.'

ggplot(data = complete_pred) +
  geom_point(aes(x = SV_length, y = weight)) +
  geom_line(aes(x = SV_length, y = general_mdl_pred), color = 'red', size = 1) +
  geom_line(aes(x = SV_length, y = whip_mdl_pred), color = 'navy', size = 1) +
  theme_minimal() +
  labs(x = 'Snout to Vent Length (mm)',
       y = 'Weight (g)',
       title = 'General NLS vs. Species Specific NLS',
      caption = str_wrap(caption)) 
```


### Data Citation
Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f


