---
title: "beepboop2"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(purrr)
library(minpack.lm)
library(broom)
library(patchwork)
library(kableExtra)
```

## 1. Overview
In this task, I will be using non linear least squares to estimate parameters of a length to weight model for lizard populations in New Mexico. 
PUT OVERVIEW HERE

## Explore the Data

### 2.1 Read in the data
```{r}
lizards <- read_csv(here("data", "lizard.csv")) %>% 
  mutate(log_length = log(SV_length), # use log() to linearize the model
         log_weight = log(weight))
```
## Set up the model

Must set up the model to be able to make guesses for coefficients a and b. Take the data and create a linear model in order for OLS regression to occur.

$$
ln(W) = ln(a) + b*ln(SLV)
$$

```{r}
lizards_mdl <- lm(log_weight ~ log_length, data = lizards) # Created for OLS

lizards_coef <- coef(lizards_mdl) %>% 
  exp() # convert coefficients out of log form

```


## Functions

The following functions are used in the analysis:

1) The model the data is being fitted to.
2) An RMSE calculation function **ALL THIS MUST BE CHANGED CUZ I PLAGERIZED**






